Combining neural networks and geostatistics for landslide hazard assessment of San Salvador Metropolitan Area, El Salvador \newline \newline Combinando redes neuronales y geoestad\'{i}stica para evaluaci\'{o}n de deslizamientos de tierra de el \'{a}rea Metropolitana de San Salvador, El Salvador

### 2015-06-04 paper.tex
03 June 2015 21:22

\author[1]{Ricardo R\'{i}os \thanks{ricardo.sv@gmail.com}}
\author[2]{Alex Rib\'{o} \thanks{alexandre4rt@gmail.com}}
\author[3]{Roberto Mej\'{i}a \thanks{robertomejia1685@gmail.com}}
\author[4]{Giovanni Molina \thanks{giova.molina@gmail.com}}

@@

\selectlanguage{spanish}
\begin{resumen}
Esta contribución describe la creaci\'{o}n de un modelo de evaluaci\'{o}n de deslizamiento de tierra para San Salvador, departamento de El Salvador. El an\'{a}lisis inicio con la obtenci\'{o}n de una foto \'{a}rea del MARN (Ministerio de Medio Ambiente y Recursos Naturales) con un total de 939407 puntos georeferenciados con el fin de producir un inventario de deslizamiento. En esta evaluaci\'{o}n de los deslizamientos se uso 4792 eventos previamente foto-interpretados y 7 factores condicionantes incluyendo: geomorfolog\'{i}a, geolog\'{i}a, precipitaciones m\'{a}ximas, aceleraciones s\'{i}smicas, pendiente del terreno, distancia a carretera y falla geol\'{o}gica. Redes Neuronales Artificiales (RNA) fueron usadas para la evaluaci\'{o}n de la susceptibilidad a deslizamiento de tierra, logrando que m\'{a}s del 80\% de deslizamientos fueran apropiadamente clasificados usando un criterio dentro y fuera de la muestra con la que se estimaron los parámetros del modelo. Regresi\'{o}n Log\'{i}stica fue usada como base de comparaci\'{o}n, obteniendo este modelo un rendimiento inferior que el de RNA con un porcentaje de correcta clasificaci\'{o}n abajo del 70\%. Para completar el an\'{a}lisis se realizo la interpolaci\'{o}n de puntos usando el m\'{e}todo krigging proveniente del enfoque geoestad\'{i}stico. Finalmente, los resultados muestran que es posible obtener un mapa de riesgo a deslizamiento de tierra, haciendo uso de una combinación de RNA y t\'{e}cnicas geoestad\'{i}sticas con lo cual la presente investigaci\'{o}n puede ayudar a la mitigaci\'{o}n de deslizamientos de tierra en El Salvador.
\end{resumen}

@@

\section{Introduction}
\label{sec:intr}
El Salvador, one of the smallest and most crowded nations in Central America, extends in Pacific coast about 240 kilometers westward from the Gulf of Fonseca to the border with Guatemala. El Salvador borders an active subduction boundary between Cocos and Caribbean plates located 30 km offshore. Therefore El Salvador is affected with high seismicity and volcanic activity related to this active boundary. There is two main sources of seismic activity, the upper interface thrust that coincides with the location of the recent volcanoes and the Benioff-Wadaty zones of subducted Cocos plate where deeper intraplate earthquakes occurs (\cite{dewey}). Due this tectonic and volcanic activity El Salvador has rugged relief. The main ranges cross the country with a rough west-east trend, parallel to the coast, these are separated from each other by faults and grabens. These ranges present several highly active volcanoes. The surface geology in El Salvador is almost entirely volcanic, dominated by upper Tertiary to Holocene volcanic rocks, only sparse outcrops of sedimentary and plutonic crops are located in the northern ranges, in the border with Honduras (\cite{weber}). Some of the most recent volcanic layers are formed by poor consolidated ashes and tuffs highly erodible (\cite{bommer}).

@@

 \begin{center}
  \begin{figure}[H]
   \centering
   \includegraphics[scale=0.80]{MASS_mapa_1}
   \caption{\small{Maps from El Salvador}}
   \label{fig:mass01}
  \end{figure}
 \end{center}

@@

\section{Landslide inventory and data sources of input variables}
\label{sec:landsinvet}
The Landslide inventory was developed from photographic analysis on the study area, using 
aerial photo from MARN which is presented in the Figure \ref{fig:img01}, where white areas represent regions of landslide occurrence which were processed using the software ILWIS.

 \begin{center}
  \begin{figure}[H]
   \centering
   \includegraphics[scale=0.40]{img01}
   \caption{\small{Aerial photo of MASS}}
   \label{fig:img01}
  \end{figure}
 \end{center}

@@

The results obtained in the geostatistics methodology showed that the spatial the data satisfies the conditions for applying krigin method. The map of landslide hazard \ref{fig:img02} was generated by krigin method and can be used by non-experts in the landslide phenomenon for many purposes such as territorial planning, prevention and mitigation of natural dissasters and so on.

>>>>

\author[1]{Ricardo R\'{i}os \thanks{ricardo.sv@gmail.com}}
\author[2]{Alexandre Rib\'{o} \thanks{alexandre4rt@gmail.com}}
\author[3]{Roberto Mej\'{i}a \thanks{robertomejia1685@gmail.com}}
\author[4]{Giovanni Molina \thanks{giova.molina@gmail.com}}

@@

\selectlanguage{spanish}
\begin{resumen}
Esta contribución describe la creaci\'{o}n de un modelo de evaluaci\'{o}n de deslizamiento de tierra para San Salvador, departamento de El Salvador. El an\'{a}lisis inicio con la obtenci\'{o}n de una foto \'{a}rea del MARN (Ministerio de Medio Ambiente y Recursos Naturales) con un total de 939407 puntos georeferenciados con el fin de producir un inventario de deslizamiento. En esta evaluaci\'{o}n de los deslizamientos se uso 4792 eventos previamente foto-interpretados y 7 factores condicionantes incluyendo: geomorfolog\'{i}a, geolog\'{i}a, precipitaciones m\'{a}ximas, aceleraciones s\'{i}smicas, pendiente del terreno, distancia a carretera y falla geol\'{o}gica. Redes Neuronales Artificiales (RNA) fueron usadas para la evaluaci\'{o}n de la susceptibilidad a deslizamiento de tierra, logrando que m\'{a}s del 80\% de deslizamientos fueran apropiadamente clasificados usando un criterio dentro y fuera de la muestra con la que se estimaron los parámetros del modelo. Regresi\'{o}n Log\'{i}stica fue usada como base de comparaci\'{o}n, obteniendo este modelo un rendimiento inferior que el de RNA con un porcentaje de correcta clasificaci\'{o}n abajo del 70\%. Para completar el an\'{a}lisis se realizo la interpolaci\'{o}n de puntos usando el m\'{e}todo kriging proveniente del enfoque geoestad\'{i}stico. Finalmente, los resultados muestran que es posible obtener un mapa de riesgo a deslizamiento de tierra, haciendo uso de una combinación de RNA y t\'{e}cnicas geoestad\'{i}sticas con lo cual la presente investigaci\'{o}n puede ayudar a la mitigaci\'{o}n de deslizamientos de tierra en El Salvador.
\end{resumen}

@@

\section{Introduction}
\label{sec:intr}
El Salvador, one of the smallest and most crowded nations in Central America, extends in Pacific coast about 240 kilometers westward from the Gulf of Fonseca to the border with Guatemala see Fig. \ref{fig:mass01}. El Salvador borders an active subduction boundary between Cocos and Caribbean plates located 30 km offshore. Therefore El Salvador is affected with high seismicity and volcanic activity related to this active boundary. There is two main sources of seismic activity, the upper interface thrust that coincides with the location of the recent volcanoes and the Benioff-Wadaty zones of subducted Cocos plate where deeper intraplate earthquakes occurs (\cite{dewey}). Due this tectonic and volcanic activity El Salvador has rugged relief. The main ranges cross the country with a rough west-east trend, parallel to the coast, these are separated from each other by faults and grabens. These ranges present several highly active volcanoes. The surface geology in El Salvador is almost entirely volcanic, dominated by upper Tertiary to Holocene volcanic rocks, only sparse outcrops of sedimentary and plutonic crops are located in the northern ranges, in the border with Honduras (\cite{weber}). Some of the most recent volcanic layers are formed by poor consolidated ashes and tuffs highly erodible (\cite{bommer}).

@@

 \begin{center}
%  \begin{figure}[H]
  \begin{figure}
   \centering
   \includegraphics[scale=0.75]{MASS_mapa_1}
   \caption{\small{Maps from El Salvador}}
   \label{fig:mass01}
  \end{figure}
 \end{center}

@@

\section{Landslide inventory and data sources of input variables}
\label{sec:landsinvet}
The Landslide inventory was developed by MARN and it was conducted in two stages. The first was carried out, based on fieldwork after the 2001 eartquakes in El Salvador. The last stage is complete with a visual inventory of places where there were landslides in visible satellite image IRSC (India) sensor of the same year.

from photographic analysis on the study area, using aerial photo from MARN which is presented in the Figure \ref{fig:img01}, where white areas represent regions of landslide occurrence which were processed using the software ILWIS.

% \begin{center}
%  \begin{figure}[H]
%   \centering
%   \includegraphics[scale=0.40]{img01}
%   \caption{\small{Aerial photo of MASS}}
%   \label{fig:img01}
%  \end{figure}
% \end{center}

@@

The results obtained in the geostatistics methodology showed that the spatial the data satisfies the conditions for applying kriging method. The map of landslide hazard \ref{fig:img02} was generated by kriging method and can be used by non-experts in the landslide phenomenon for many purposes such as territorial planning, prevention and mitigation of natural dissasters and so on.
